---
title: "Song songdata"
author: "Silvia Ventoruzzo"
date: "10/5/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Preparation

Packages & functions
```{r}
library(tidyverse)
library(tidytext)
library(DataExplorer)
library(gridExtra)
library(grid)
library(tm)
library(topicmodels)
library(gtools)
library(gtable)
library(quanteda)
library(stopwords)
library(reshape2)
library(ldatuning)
library(cld3)
library(SnowballC)
library(xtable)
```

Load dataset
```{r}
songs_genre <- read_csv("./Data/lyrics.csv.zip")
songs       <- read_csv("./Data/songdata.csv.zip")
```

Join datasets
```{r}
# Clean before joining
songs_genre <- songs_genre %>%
  dplyr::select(-index, -lyrics) %>%
  mutate(song   = gsub("-", " ", song),
         artist = gsub("-", " ", artist),
         artist = ifelse(artist == "beyonce knowles", "beyonce", artist)) %>%
  filter(genre != "Not Available") %>%
  mutate_if(is.character, tolower)

songs <- songs %>%
  dplyr::select(-link) %>%
  filter(!is.na(text)) %>%
  mutate_if(!(grepl("text", names(.))), tolower)

# Join
songdata <- songs %>%
  inner_join(songs_genre, by = c("artist", "song"))

rm("songs_genre", "songs")
```

We would like to keep only 200 songs, but at the moment we have 11555. We initially look at the language, restricting our dataset of only songs in English, and length of the song, keeping only songs with length bigger than 1st quantile.
```{r}
# Language
songdata <- songdata %>%
  mutate(text_language = detect_language(text))

songdata %>%
  count(text_language) %>%
  mutate(text_language = reorder(text_language, -n)) %>%
  ggplot() +
  geom_bar(aes(x = text_language, y = n), stat = "identity")

songdata <- songdata %>%
  filter(text_language == "en") %>%
  dplyr::select(-text_language)

# Length
songdata <- songdata %>%
  mutate(text_length = str_length(songdata$text))

songdata %>%
  ggplot() +
  geom_boxplot(aes(x = factor(0), y = text_length)) +
  coord_flip()

songdata <- songdata %>%
  filter(text_length > quantile(text_length, 0.25)) %>%
  dplyr::select(-text_length)
```

We now look at the distributions of the variables artist, year and genre.
```{r}
# Year
songdata %>%
  count(year) %>%
  mutate(occurences = n/sum(n)) %>%
  ggplot() +
  geom_bar(aes(x = year, y = occurences), stat = "identity") +
  scale_x_continuous(breaks = seq(min(songdata$year), max(songdata$year), 5)) +
  scale_y_continuous(labels = scales::percent_format()) +
  labs(title    = "Distribution of song years",
       subtitle = "Initial data")

# Genre
songdata %>%
  count(genre) %>%
  mutate(occurences = n/sum(n)) %>%
  ggplot() +
  geom_bar(aes(x = reorder(genre, -occurences), y = occurences), stat = "identity") +
  scale_y_continuous(labels = scales::percent_format()) +
  labs(x        = "genre",
       title    = "Distribution of music genres",
       subtitle = "Initial data")

# Artist
songdata %>%
  count(artist) %>%
  mutate(occurences = n/sum(n)) %>%
  ggplot() +
  geom_bar(aes(x = reorder(artist, occurences), y = occurences), stat = "identity") +
  scale_y_continuous(labels = scales::percent_format()) +
  coord_flip() +
  labs(x        = "artist",
       title    = "Distribution of artists",
       subtitle = "Initial data")
```

We can see that great part of the songs are from 2005 onwards. We will therefore firstly restrict the dataset to the songs from 2010. We will furthermore only keep songs of the 5 most common music genres (pop, rock, hip-hop, jazz, country).
```{r}
songdata <- songdata %>%
  filter(year >= 2010,
         genre %in% c("pop", "rock", "hip-hop", "jazz", "country"))
```

We still have too many songs (1483). We will thus now look at artists. We wish to create some variability in the texts, therefore it would be good to have a mixture of different artists. I would however prefer having famous artists, to make the project more relatable, but also considering a mix of genres. Following artists will be kept:
- adele
- aerosmith
- avril lavigne
- backstreet boys
- billie holiday
- bon jovi
- coldplay
- depeche mode
- dolly parton
- drake
- ed sheeran
- ella fitzgerald
- eminem
- enrique iglesias
- evanescence
- foo fighters
- fall out boy
- george jones
```{r}
set.seed(62781649)
songdata <- songdata %>%
  filter(artist %in% c("adele", "aerosmith", "avril lavigne", "backstreet boys", "billie holiday", "bon jovi", "coldplay", "depeche mode", "dolly parton", "drake", "ed sheeran", "ella fitzgerald", "eminem", "evanescence", "foo fighters", "fall out boy", "george jones")) %>%
  sample_n(200)
```

Produce now final plots
```{r}
genre <- songdata %>%
  count(genre) %>%
  ungroup() %>%
  rename(level = genre) %>%
  mutate(occurences = n/sum(n),
         level      = reorder(level, occurences),
         variable   = "genre")
artist <- songdata %>%
  count(artist) %>%
  ungroup() %>%
  rename(level = artist) %>%
  mutate(occurences = n/sum(n),
         level      = reorder(level, occurences),
         variable   = "artist")
year <- songdata %>%
  count(year) %>%
  ungroup() %>%
  rename(level = year) %>%
  mutate(occurences = n/sum(n),
         level      = reorder(level, -level),
         variable   = "year")
genre %>%
  rbind(artist) %>%
  rbind(year) %>%
  ggplot() +
  # geom_bar(aes(x = reorder(artist, occurences), y = occurences), stat = "identity", fill = "chartreuse4") +
  geom_bar(aes(x = level, y = occurences, fill = variable), stat = "identity") +
  theme_bw() +
  guides(fill = FALSE) +
  scale_y_continuous(labels = scales::percent_format(accuracy = 1)) +
  scale_fill_manual(values = c("chartreuse3", "#E69F00", "#56B4E9")) +
  labs(y = "Percentage frequency",
       x = "Values") +
  theme(axis.text.y  = element_text(size = 10),
        axis.text.x  = element_text(size = 10),
        strip.text.x = element_text(size = 11)) +
  coord_flip() +
  facet_wrap(~variable, scales = "free", nrow = 3)
rm("genre", "artist", "year")
```

Latex table
```{r}
songdata %>%
  arrange(genre, artist) %>%
  select(genre, artist, song, year) %>%
  mutate(year = as.integer(year)) %>%
  xtable() %>%
  print(include.rownames = FALSE)
```

Produce plot of genres for poster
```{r}
genre <- songdata %>%
  count(genre) %>%
  mutate(occurences = n/sum(n),
         variable   = "genre") %>%
  rename(level = genre)
artist <- songdata %>%
  count(artist) %>%
  mutate(occurences = n/sum(n),
         variable   = "artist") %>%
  rename(level = artist)
genre %>%
  rbind(artist) %>%
  ggplot() +
  # geom_bar(aes(x = reorder(artist, occurences), y = occurences), stat = "identity", fill = "chartreuse4") +
  geom_bar(aes(x = reorder(level, occurences), y = occurences, fill = level), stat = "identity") +
  guides(fill = FALSE) +
  scale_y_continuous(labels = scales::percent_format(accuracy = 1)) +
  theme(axis.title.y = element_blank(),
        axis.text.y  = element_text(size = 14),
        axis.text.x  = element_text(size = 14),
        axis.title.x = element_blank(),
        strip.text.x = element_text(size = 14)) +
  coord_flip() +
  facet_wrap(~variable, scales = "free")
rm("genre", "artist")
```

Look at summed percentages
```{r}
# Artist
songdata %>%
  count(artist) %>%
  arrange(-n) %>%
  mutate(perc = n/sum(n), 
         cumsum = cumsum(perc))

# Genre
songdata %>%
  count(genre) %>%
  arrange(n) %>%
  mutate(perc = n/sum(n), 
         cumsum = cumsum(perc))
```

Combination artist-genre
Each artist is associated to only one genre
```{r}
songdata %>%
  count(artist, genre) %>%
  count(artist)
```

Text length
```{r}
# songdata %>%
#   mutate(text_length = str_length(songdata$text)) %>%
#   ggplot() +
#   geom_boxplot(aes(x = factor(0), y = text_length, fill = genre)) +
#   guides(fill = FALSE) +
#   theme(axis.title.y = element_blank(),
#         axis.text.y  = element_blank(),
#         axis.ticks.y = element_blank(),
#         axis.title.x = element_text(size = 12)) +
#   labs(y = "text length") +      
#   coord_flip() +
#   facet_wrap(~genre)
```

text
```{r}
# songdata <- songdata %>%
#   mutate(text = tolower(text),
#          text = removePunctuation(text),
#          text = removeNumbers(text),
#          text = stripWhitespace(text),
#          text = stemDocument(text))
```

genre - year
```{r}
songdata %>%
  count(year, genre) %>%
  group_by(year) %>%
  mutate(percentage = n/sum(n)) %>%
  ungroup() %>%
  ggplot() +
  geom_line(aes(x = year, y = percentage, color = genre))

# # From: https://towardsdatascience.com/text-analytics-topic-modelling-on-music-genres-song-songdata-deb82c86caa2
# songdata %>% 
#   mutate(fiveyears = as.character(year) %>%
#                       paste("01", "01", sep = "-") %>%
#                       as.Date() %>%
#                       lubridate::floor_date("y year")) %>%
#   count(fiveyears, genre) %>% 
#   group_by(fiveyears) %>%
#   mutate(freq = round(n/sum(n), 2)) %>% 
#   # filter(genre %in% c("Country", "Hip-Hop", "Metal", "Pop", "Rock")) %>% 
#   ggplot(aes(fiveyears, freq, colour = genre)) +
#   # geom_line() +
#   geom_smooth(se = FALSE) +
#   labs(x = "year", y = "smoothed")+
#   scale_y_continuous(labels = scales::percent_format()) 
```

# Data preparation

Stopwords
```{r}
# Standard
custom_stopwords <- lapply(stopwords_getsources()[stopwords_getsources() != "misc"], 
                    function(x) stopwords::stopwords(language = "en", source = x)) %>%
  unlist()

# Custom
custom_stopwords <- c(custom_stopwords,
               "ooh", "tat", "yeah", "du", "wanna", "gonna", "uh", "gotta", "hey", "chorus",
               "ya", "verse", "m-my", "gon", "echo", "hook", "ding", "dong", "yi", "ti", "comma")
```

For Latex
```{r}
c("ooh", "tat", "yeah", "du", "wanna", "gonna", "uh", "gotta", "hey", "chorus",
               "ya", "verse", "m-my", "gon", "echo", "hook", "ding", "dong", "yi", "ti", "comma") %>%
  as.data.frame() %>%
  rename(Words = ".") %>%
  arrange(Words) %>%
  xtable() %>%
  print(include.rownames = FALSE)
```

Tokenization for analysis
```{r}
songdata_words <- songdata %>%
  unnest_tokens(output = "word", input = "text", token = "words") %>%
  filter(!(word %in% custom_stopwords))
```

Number of words
```{r}
songdata_words %>%
  count(genre, song) %>%
  ggplot() +
  geom_boxplot(aes(x = genre, y = n, fill = genre, color = genre), alpha = 0.5) +
  theme_bw() +
  guides(fill = FALSE, color = FALSE) +
  labs(y = "Number of words",
       x = "Genre") +
  theme(axis.text.y  = element_text(size = 10),
        axis.text.x  = element_text(size = 10)) +
  coord_flip()
```

Number of distinct words
```{r}
songdata_words %>%
  group_by(genre, song) %>%
  summarize(n = n_distinct(word)) %>%
  ggplot() +
  geom_boxplot(aes(x = genre, y = n, fill = genre, color = genre), alpha = 0.5) +
  theme_bw() +
  guides(fill = FALSE, color = FALSE) +
  labs(y = "Number of distinct words",
       x = "Genre") +
  theme(axis.text.y  = element_text(size = 10),
        axis.text.x  = element_text(size = 10)) +
  coord_flip()
```

```{r}
songdata_words %>%
  group_by(genre, song) %>%
  summarize(total    = n(),
            distinct = n_distinct(word)) %>%
  ungroup() %>%
  gather(key = "words", value = "count", total, distinct) %>%
  ggplot() +
  geom_boxplot(aes(x = words, y = count, fill = genre, color = genre), alpha = 0.5) +
  theme_bw() +
  guides(fill = FALSE, color = FALSE) +
  labs(y = "Number of words",
       x = "Genre") +
  theme(axis.text.y  = element_text(size = 10),
        axis.text.x  = element_text(size = 10),
        strip.text.x = element_text(size = 11)) +
  coord_flip() +
  facet_wrap(~genre, ncol = 1)
```


Most common words per genre
```{r}
songdata_words %>%
  mutate(word = wordStem(word, language = "english"),
         word = ifelse(word == "shit", "sh*t",
                ifelse(word == "fuck", "f*ck",
                ifelse(word == "bitch", "b*tch",
                ifelse(word == "nigga", "n*gga", word))))) %>%
  count(genre, word) %>%
  group_by(genre) %>%
  top_n(10, n) %>%
  ungroup() %>%
  group_by(genre) %>%
  mutate(percentage = n/sum(n)) %>%
  ungroup() %>%
  # mutate(word  = reorder(word, n)) %>%
  ggplot(aes(reorder(word, percentage), percentage, fill = factor(genre))) +
  geom_col(show.legend = FALSE) +
  theme_bw() +
  labs(x = "Most common words",
       y = "Percentage frequency") +
  facet_wrap(~ genre, scales = "free_y") +
  theme(axis.text.x  = element_text(size = 10),
        axis.text.y  = element_text(size = 10),
        strip.text.x = element_text(size = 11)) +
  scale_y_continuous(labels = scales::percent_format(accuracy = 1)) +
  coord_flip()
```

# Sentiment analysis

Plot sentiments by genre
NEED TO FIND THE RIGHT WAY TO WEIGHT THE SENTIMENT
```{r}
songdata_words %>%
  group_by(genre, song) %>%
  summarize(n = n_distinct(word))

songdata_words %>% 
  inner_join(get_sentiments("afinn"), by = "word") %>% 
  group_by(genre, song) %>% 
  summarize(n                  = n(),
            # n                  = n_distinct(word),
            sentiment          = sum(score),
            weighted_sentiment = sentiment/n) %>% 
  ungroup() %>%
  mutate(genre = reorder(genre, desc(genre))) %>%
  ggplot() +
  geom_boxplot(aes(x = genre, y = weighted_sentiment, fill = reorder(genre, desc(genre)), color = reorder(genre, desc(genre))),
               alpha = 0.5) +
  theme_bw() +
  guides(fill = FALSE, color = FALSE) +
  labs(x = "Genre",
       y = "Total sentiment divided by total number of words") +
  theme(axis.text.x  = element_text(size = 10),
        axis.text.y  = element_text(size = 10)) +
  coord_flip()

# bing_and_nrc <- bind_rows(
#   songdata_words %>% 
#     inner_join(get_sentiments("bing"), by = "word") %>%
#     mutate(method = "BING"),
#   songdata_words %>% 
#     inner_join(get_sentiments("nrc") %>% filter(sentiment %in% c("positive", "negative")),
#                by = "word") %>%
#     mutate(method = "NRC")) %>%
#   count(method, genre, sentiment) %>%
#   spread(sentiment, n, fill = 0) %>%
#   mutate(sentiment = positive - negative)
# 
# bind_rows(afinn, 
#           bing_and_nrc) %>%
#   ggplot(aes(genre, sentiment, fill = method)) +
#   geom_col(show.legend = FALSE) +
#   coord_flip() +
#   facet_wrap(~method, ncol = 1) +
#   theme(axis.title.x = element_text(size = 12),
#         axis.title.y = element_blank())
```

Plot only one sentiment by genre for Poster
```{r}
songdata_words %>% 
  inner_join(get_sentiments("afinn"), by = "word") %>% 
  group_by(genre) %>% 
  summarise(score = sum(score)) %>%
  mutate(sentiment = "sentiment") %>%
  ggplot(aes(x = genre, y = score)) +
  geom_bar(stat = "identity", fill = "dodgerblue") +
  coord_flip() +
  theme(axis.title.x = element_blank(),
        axis.text.x  = element_text(size = 10),
        axis.text.y  = element_text(size = 10),
        axis.title.y = element_blank()) +
  facet_wrap(~sentiment)

songdata_words %>% 
  inner_join(get_sentiments("afinn"), by = "word") %>% 
  group_by(genre, song) %>% 
  summarise(score = sum(score)) %>%
  mutate(sentiment = "sentiment") %>%
  ggplot(aes(x = genre, y = score)) +
  geom_boxplot(aes(fill = genre, color = genre), alpha = 0.5) +
  coord_flip() +
  guides(fill = FALSE, color = FALSE) +
  theme(axis.title.x = element_blank(),
        axis.text.x  = element_text(size = 14),
        axis.text.y  = element_text(size = 14),
        strip.text.x = element_text(size = 14),
        axis.title.y = element_blank()) +
  facet_wrap(~sentiment)
```

Plot sentiments by artist
```{r}
songdata_words %>% 
  inner_join(get_sentiments("afinn"), by = "word") %>% 
  group_by(genre, artist, song) %>% 
  summarize(n                  = n(),
            # n                  = n_distinct(word),
            sentiment          = sum(score),
            weighted_sentiment = sentiment/n) %>% 
  ggplot() +
  geom_boxplot(aes(x = artist, y = weighted_sentiment, fill = genre, color = genre), alpha = 0.5) +
  theme_bw() +
  guides(fill = FALSE, color = FALSE) +
  labs(x = "Genre and corresponding artists",
       y = "Total sentiment divided by total number of words") +
  theme(axis.text.x  = element_text(size = 10),
        axis.text.y  = element_text(size = 10)) +
  coord_flip() +
  facet_grid(genre ~ ., scales = "free_y", switch = "y")
```


# TF-IDF

```{r}
# songdata_tf_idf <- songdata_words %>%
#   count(genre, artist, song, word, sort = TRUE) %>%
#   # mutate(document = paste(genre, artist, song, sep = "_")) %>%
#   bind_tf_idf(word, song, n)
```

High tf-idf words in genres
```{r}
# songdata_tf_idf %>%
#   arrange(desc(tf_idf)) %>%
#   mutate(word = factor(word, levels = rev(unique(word)))) %>% 
#   group_by(genre) %>% 
#   top_n(20) %>% 
#   ungroup() %>%
#   ggplot(aes(word, tf_idf, fill = genre)) +
#   geom_col(show.legend = FALSE) +
#   labs(x = NULL, y = "tf-idf") +
#   facet_wrap(~genre, scales = "free") +
#   coord_flip()
```

Distribution of tf-idf
```{r}
# songdata_tf_idf %>%
#   ggplot() +
#   geom_boxplot(aes(x = factor(0), y = tf_idf)) +
#   coord_flip()
```

# Find best number of topics

Tokenize
```{r}
# songdata_dtm <- songdata_words %>%
#   count(artist, song, word, sort = TRUE) %>%
#   ungroup() %>%
#   cast_dtm(term = "word", document = "song", value = "n")

songdata_tokens <- songdata %>%
  corpus() %>%
  tokens(what = "word",
         remove_numbers = TRUE, remove_punct = TRUE,
         remove_symbols = TRUE, remove_separators = TRUE)

docnames(songdata_tokens) <- paste(songdata$artist, songdata$song, sep = ";")
```

Remove words that have an tf-idf lower than 1st quantile
```{r}
# songdata_words <- songdata_tf_idf %>%
#   filter(tf_idf > quantile(tf_idf, probs = 0.25)) %>%
#   inner_join(songdata_words, by = c("genre", "artist", "song", "word"))
# 
# test <- songdata_words %>%
#   mutate(document = paste(artist, song, sep = "_")) %>%
#   pull(word) %>%
#   dfm(tolower = TRUE, stem = TRUE, remove = custom_stopwords) %>%
#   convert(to = "tm") %>%
#   TermDocumentMatrix()
# 
# test$dimnames$Docs <- paste(songdata$artist, songdata$song, sep = ";")
```

Transform all dataset into DocumentTermMatrix
```{r}
songdata_dtm <- songdata_tokens %>%
  dfm(tolower = TRUE, stem = TRUE, remove = custom_stopwords) %>%
  convert(to = "tm")

rm("songdata_tokens")
```

Perplexity for hold-out set
```{r}
# Train DTM
set.seed(4324663) 
sampling <- sample(1:nrow(songdata), replace = FALSE,size = nrow(songdata)*0.7)
train_data <- songdata[sampling,]
DTM_train <- train_data %>%
  corpus() %>%
  tokens(what = "word",
         remove_numbers = TRUE, remove_punct = TRUE,
         remove_symbols = TRUE, remove_separators = TRUE)
docnames(DTM_train) <- paste(train_data$artist, train_data$song, sep = ";")
DTM_train <- DTM_train %>%
  dfm(tolower = TRUE, stem = TRUE, remove = custom_stopwords) %>%
  convert(to = "tm")

# Test DTM
test_data <- songdata[-sampling,]
DTM_test <- test_data %>%
  corpus() %>%
  tokens(what = "word",
         remove_numbers = TRUE, remove_punct = TRUE,
         remove_symbols = TRUE, remove_separators = TRUE)
docnames(DTM_test) <- paste(test_data$artist, test_data$song, sep = ";")
DTM_test <- DTM_test %>%
  dfm(tolower = TRUE, stem = TRUE, remove = custom_stopwords) %>%
  convert(to = "tm")

rm("sampling")

perplexity_df <- data.frame(topics = numeric(), train = numeric(), test = numeric())
topics <- c(2:200)

# set.seed(4324663)
for (i in topics) {
  cat(paste(i, "topics..."))
  fitted <- LDA(DTM_train, k = i, method = "VEM",
                control = list(seed = 4324663))
  perplexity_df[i-1,1]  <- i
  perplexity_df[i-1,2]  <- perplexity(fitted, newdata = DTM_train)
  perplexity_df[i-1,3]  <- perplexity(fitted, newdata = DTM_test)
  cat(paste("done", "\n"))
}

rm("DTM_test", "DTM_train", "fitted", "test_data", "train_data", "i", "topics", "custom_stopwords", "sampling")
```

Rest of metrics
```{r}
other_metrics <- FindTopicsNumber(dtm     = songdata_dtm,
                                  topics  = seq(2, 200, by = 1),
                                  metrics = c("Griffiths2004", "CaoJuan2009", "Arun2010", "Deveaud2014"),
                                  method  = "VEM",
                                  control = list(seed = 4324663),
                                  verbose = TRUE)

# save.image(file = "k_metrics.RData")
```

Join
```{r}
k_metrics <- perplexity_df %>%
  rename(Perplexity = test) %>%
  select(-train) %>%
  inner_join(other_metrics, by = "topics")

rm(perplexity_df)
rm(other_metrics)
```

Plot
Plot of metrics for Poster
```{r}
k_metrics %>%
  mutate_if(!grepl("topics", names(.)), function(x) 
                            ifelse(is.infinite(x), NA, 
                                   scales::rescale(x, to = c(0, 1), from = range(x, na.rm = TRUE, finite = TRUE)))) %>%
  gather(key = "metric", value = "value", Perplexity, CaoJuan2009, Arun2010, Deveaud2014) %>%
  mutate(optimization = ifelse(metric == "Deveaud2014", "maximize", "minimize")) %>%
  ggplot(aes(x = topics, y = value, color = metric)) +
  geom_line(stat = "identity") +
  geom_point(aes(shape = metric, fill = metric), size = 1) +
  theme_bw() +
  scale_shape_manual(values = 21:24) +
  scale_x_continuous(breaks = seq(0, max(k_metrics$topics), 25)) +
  scale_y_continuous(oob = scales::discard) +
  labs(x = "Number of topics",
       y = "Scaled values of metrics") +
  theme(legend.position   = "top",
        legend.margin     = margin(0,0,0,0),
        # legend.box.margin = margin(-5,-5,-5,-5),
        # legend.title      = element_blank(),
        axis.text.x       = element_text(size = 10),
        strip.text.y      = element_text(size = 11)) +
  facet_grid(optimization ~ .)
```

Find optimal number for each metric
```{r}
metrics <- k_metrics %>%
  select_if(!grepl("topics", names(.))) %>%
  colnames()

k_optimal <- data.frame(metrics          = metrics,
                        number_of_topics = rep(NA, 4))

for (m in metrics) {
  
  cat(paste("Metric: ", m, ", "))

  if (m == "Deveaud2014") {
    
    optimal_value <- k_metrics %>%
      mutate(Deveaud2014 = ifelse(is.infinite(Deveaud2014), NA, Deveaud2014)) %>%
      pull(!!m) %>%
      max(na.rm = TRUE)
  
  } else {
    
    optimal_value <- k_metrics %>%
      pull(!!m) %>%
      min(na.rm = TRUE)

  }
  
  cat(paste("Optimal value: ", optimal_value, ", "))
  
  optimal_k <- k_metrics %>%
    filter(!!sym(m) == optimal_value) %>%
    pull(topics)
  
  cat(paste("Optimal number of topics: ", optimal_k, "\n"))
  
  k_optimal <- k_optimal %>%
    mutate(number_of_topics = ifelse(metrics == m, optimal_k, number_of_topics))
  
}

k_optimal %>%
  xtable()

rm(metrics)
rm(m)
rm(optimal_value)
rm(optimal_k)
```

Plot of metrics for Poster
```{r}
k_metrics %>%
  filter(topics %in% 2:20) %>%
  mutate_if(!grepl("topics", names(.)), function(x) 
                            scales::rescale(x, to = c(0, 1), from = range(x, na.rm = TRUE, finite = TRUE))) %>%
  gather(key = "metric", value = "value", Perplexity, CaoJuan2009, Arun2010, Deveaud2014) %>%
  mutate(optimization = ifelse(metric == "Deveaud2014", "maximize", "minimize")) %>%
  ggplot(aes(x = topics, y = value, color = metric)) +
  geom_line(stat = "identity") +
  geom_point(aes(shape = metric, fill = metric), size = 3) +
  scale_shape_manual(values = 21:24) +
  scale_x_continuous(breaks = k_metrics$topics) +
  theme(legend.position   = "top",
        legend.margin     = margin(0,0,0,0),
        legend.box.margin = margin(-5,-5,-5,-5),
        legend.title      = element_blank(),
        axis.title.x      = element_blank(),
        axis.ticks.y      = element_blank(),
        axis.text.x       = element_text(size = 14),
        axis.text.y       = element_blank(),
        strip.text.y      = element_text(size = 14),
        axis.title.y      = element_blank()) +
  facet_grid(optimization ~ .)
```

LDA with values of k from 2 to 5

```{r}
songdata_lda <- list()

for (k in 2:5) {
  cat(paste("LDA with k =", k, "..."))
  songdata_lda[[k-1]] <- LDA(songdata_dtm, k = k, method = "VEM", control = list(seed = 978324))
  cat("done\n")
}
names(songdata_lda) <- c(2:(length(songdata_lda)+1))

rm(k)

# save.image(file = "lda_25k.RData")
# 
# load(file = "lda_25k.RData")
```

Calculate probabilites of interest:
- beta = per-topic-per-word probabilities
- lambda = per-document-per-topic probabilities
```{r}
# beta
songdata_beta <- list()
for (k in 2:(length(songdata_lda)+1)) {
  songdata_beta[[k-1]] <- tidy(songdata_lda[[k-1]], matrix = "beta") %>%
    mutate(topics = k)
}
names(songdata_beta) <- c(2:(length(songdata_beta)+1))

# gamma
songdata_gamma <- list()
for (k in 2:(length(songdata_lda)+1)) {
  songdata_gamma[[k-1]] <- tidy(songdata_lda[[k-1]], matrix = "gamma") %>%
    mutate(topics = k)
}
names(songdata_gamma) <- c(2:(length(songdata_gamma)+1))

rm(k)
```

```{r}
# test_lda <- LDA(songdata_dtm, k = 3, method = "VEM", control = list(seed = 978324))
# 
# test_beta <- tidy(test_lda, matrix = "beta")
# 
# test_gamma <- tidy(test_lda, matrix = "gamma")
```

Words in topics
```{r}
for (k in 2:(length(songdata_beta)+1)) {
  plot <- songdata_beta[[k-1]] %>%
    group_by(topic) %>%
    top_n(10, beta) %>%
    mutate(term = ifelse(term == "shit", "sh*t",
                  ifelse(term == "fuck", "f*ck",
                  ifelse(term == "bitch", "b*tch",
                  ifelse(term == "nigga", "n*gga", term))))) %>%
    ungroup() %>%
    arrange(topic, -beta) %>%
    mutate(term  = reorder(term, topic),
           topic = paste0("topic", topic)) %>%
    ggplot(aes(reorder(term, beta), beta, fill = factor(topic))) +
    geom_col(show.legend = FALSE) +
    theme_bw() +
    labs(x = "Words",
         y = "Probability") +
    facet_wrap(~ topic, scales = "free_y") +
    theme(axis.text.x = element_text(size = 10),
          axis.text.y = element_text(size = 10)) +
    coord_flip()
  
  print(plot)
}

rm(plot)
rm(k)
```

Words in topics (3 topics) for Poster
```{r}
songdata_beta[["3"]] %>%
    group_by(topic) %>%
    top_n(20, beta) %>%
    mutate(term = ifelse(term == "shit", "sh*t",
                  ifelse(term == "fuck", "f*ck",
                  ifelse(term == "bitch", "b*tch",
                  ifelse(term == "nigga", "n*gga", term))))) %>%
    ungroup() %>%
    arrange(topic, -beta) %>%
    mutate(term  = reorder(term, topic),
           topic = paste0("topic", topic)) %>%
    ggplot(aes(reorder(term, beta), beta, fill = factor(topic))) +
    geom_col(show.legend = FALSE) +
    facet_wrap(~ topic, scales = "free_y") +
    theme(axis.title.x      = element_blank(),
          axis.text.x       = element_text(size = 10),
          axis.text.y       = element_text(size = 13),
          strip.text.x      = element_text(size = 14),
          axis.title.y      = element_blank()) +
    coord_flip()
```

Weighted average sentiment pro topic
```{r}
do.call("rbind", songdata_beta) %>%
  inner_join(get_sentiments("afinn"), by = c("term" = "word")) %>% 
  mutate(topic          = paste0("topic", topic),
         weighted_score = beta*score,
         topics         = paste(topics, "topics")) %>%
  group_by(topics, topic) %>% 
  summarize(sentiment = sum(weighted_score)) %>%
  ggplot(aes(topic, sentiment, fill = topic)) +
  geom_col(show.legend = FALSE) +
  theme_bw() +
  theme(axis.text.y  = element_text(size = 10),
        axis.text.x  = element_text(size = 10),
        strip.text.x = element_text(size = 11)) +
  labs(y = "Weighted average sentiment",
       x = "Topics") + 
  facet_wrap(~topics, scales = "free_x")
```

Weighted average sentiment pro topic (3 topics) for Poster
```{r}
songdata_beta[["3"]] %>%
  inner_join(get_sentiments("afinn"), by = c("term" = "word")) %>% 
  mutate(topic          = paste0("topic", topic),
         weighted_score = beta*score,
         topics         = paste(topics, "topics")) %>%
  group_by(topics, topic) %>% 
  summarize(sentiment = sum(weighted_score)) %>%
  ggplot(aes(topic, sentiment, fill = topic)) +
  geom_col(show.legend = FALSE) +
  theme_bw() +
  theme(axis.text.x = element_text(size = 10),
        axis.text.y = element_text(size = 10)) +
  labs(y = "Weighted average sentiment",
       x = "Topics") +
  coord_flip()
```

Gamma for genres
```{r}
for (k in 2:(length(songdata_gamma)+1)) {
  
  genre_gamma <- songdata_gamma[[k-1]] %>%
    separate(document, into = c("artist", "song"), sep = ";") %>%
    inner_join(songdata, by = c("artist", "song")) %>%
    dplyr::select(-year, -text)
  
  plot <- genre_gamma %>%
    mutate(genre = reorder(genre, gamma * topic),
           topic = paste0("topic", topic)) %>%
    ggplot(aes(factor(topic), gamma, fill = factor(topic), color = factor(topic))) +
    geom_boxplot(alpha = 0.5) +
    theme_bw() +
    theme(axis.text.x  = element_text(size = 10),
          axis.text.y  = element_text(size = 10),
          strip.text.x = element_text(size = 11)) +
    guides(fill = FALSE, color = FALSE) +
    labs(x = "Topics",
         y = "Probability") +
    facet_wrap(~ genre)
  
  print(plot)
}

rm(genre_gamma)
rm(plot)
rm(k)
```

Idea: for each document only keep the topic where it has the highest probability

Gamma for years
```{r}
for (k in 2:(length(songdata_gamma)+1)) {
  
  year_gamma <- songdata_gamma[[k-1]] %>%
    separate(document, into = c("artist", "song"), sep = ";") %>%
    inner_join(songdata, by = c("artist", "song")) %>%
    dplyr::select(-text)
  
  plot <- year_gamma %>%
    mutate(year = reorder(year, gamma * topic),
           topic = paste0("topic", topic)) %>%
    ggplot(aes(factor(topic), gamma, fill = factor(topic), color = factor(topic))) +
    geom_boxplot(alpha = 0.5) +
    theme_bw() +
    labs(y = "Probability",
         x = "Topics") +
    theme(axis.title.x = element_blank()) +
    guides(fill = FALSE, color = FALSE) +
    facet_wrap(~year)
  
  print(plot)
}

rm(year_gamma)
rm(plot)
```

Gamma for artists
```{r}
for (k in 2:(length(songdata_gamma)+1)) {

  plot <- songdata_gamma[[k-1]] %>%
    separate(document, into = c("artist", "song"), sep = ";") %>%
    inner_join(songdata, by = c("artist", "song")) %>%
    dplyr::select(-year, -text) %>%
    mutate(topic = paste0("topic", topic)) %>%
    ggplot() +
    geom_boxplot(aes(x = artist, y = gamma, fill = topic, color = topic), alpha = 0.5) +
    theme_bw() +
    guides(fill = FALSE, color = FALSE) +
    labs(x = "Artists",
         y = "Probability") +
    theme(axis.text.x   = element_text(size = 10, angle = 45, hjust = 0.8, vjust = 1),
          axis.text.y   = element_text(size = 10)) +
    coord_flip() +
    facet_grid(genre ~ topic, scales = "free_y", switch = "y")

        print(plot)

}

rm(plot)
rm(k)
```

```{r}
# for (k in 2:(length(songdata_gamma)+1)) {
  
for (k in 2:2) {
  
  plot <- songdata_gamma[[k-1]] %>%
    separate(document, into = c("artist", "song"), sep = ";") %>%
    inner_join(songdata, by = c("artist", "song")) %>%
    dplyr::select(-year, -text) %>%
    mutate(topic = paste0("topic", topic)) %>%
    ggplot() +
    # geom_col(aes(artist, gamma, fill = topic), show.legend = FALSE) +
    geom_boxplot(aes(factor(topic), gamma, fill = factor(topic), color = factor(topic)), alpha = 0.5, show.legend = FALSE) +
    theme_bw() +
    # coord_flip() +
    # facet_grid(rows = vars(topic), cols = vars(genre), scales = "free_x", switch = "x") +
    facet_wrap(. ~ genre + artist) +
    labs(x = "Topics",
         y = "Probability") +
    theme(axis.text.x  = element_text(size = 10),
          axis.text.y  = element_text(size = 9),
          strip.text.x = element_text(size = 10))
  
  print(plot)
  
}

rm(k)
rm(plot)
```

Top 20 words in topics
```{r}
for (k in 2:(length(songdata_beta)+1)) {
  plot <- songdata_beta[[k-1]] %>%
    group_by(topic) %>%
    top_n(20, beta) %>%
    mutate(term = ifelse(term == "shit", "sh*t",
                  ifelse(term == "fuck", "f*ck",
                  ifelse(term == "bitch", "b*tch",
                  ifelse(term == "nigga", "n*gga", term))))) %>%
    ungroup() %>%
    arrange(topic, -beta) %>%
    mutate(term  = reorder(term, topic),
           topic = paste0("topic", topic)) %>%
    ggplot(aes(reorder(term, beta), beta, fill = factor(topic))) +
    geom_col(show.legend = FALSE) +
    labs(x = "terms") +
    facet_wrap(~ topic, scales = "free_y") +
    theme(axis.title.x = element_blank()) +
    coord_flip()
  
  print(plot)
}

rm(plot)
rm(k)
```